Offline Reinforcement Learning with Tsallis Regularization

### 2023-04-18 paper/main.tex
updated the paper with KL divergence upper bound

A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
\begin{align}
    \begin{split}
        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\frac{1}{2}\AdaAngleProduct{\pi}{\logq{\pi}}.
    \end{split}
\end{align}
When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
\begin{align*}
    \pi_{k+1}(a|s) = &\AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau q} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau q}}}, \\
    &\qquad \qquad \psi\AdaBracket{\frac{Q_{\pi}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|},
\end{align*}
where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \includegraphics[width=\textwidth]{img/scaled_Gaussian_sparsemax.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.475\textwidth}
        \includegraphics[width=\textwidth]{img/scaled_Boltzmann_sparsemax.png}
    \end{subfigure}
    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies by truncating actions with values lower than $\psi$. 
    The truncated actions can be seen as not present in the offline dataset.
    }
\end{figure}

@@

The key to both in-sample max \cite{Fujimoto2019-InSampleMax} and in-sample softmax \cite{Xiao2023-InSampleSoftmax} is the dataset support constraint.
In this paper, we propose to exploit the properties of sparsemax policies to naturally satisfy this constraint.
In a nutshell, we assume the behavior policies are Tsallis sparsemax policies, and therefore the actions not present in the dataset are with low probability and are exactly these truncated actions.
Formally, suppose the dataset is generated by:
\begin{align}
    \datasetPolicy(a|s) = \exp_2\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
\end{align}
where $\tau_\mathcal{D}$ is a unknown coefficient and $K_\mathcal{D}(s)$ denotes the set of allowable actions.
Now let us inspect the Bellman equation for the behavior policy:
\begin{align}
    \begin{split}
    Q_{\datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \datasetPolicy(a'|s') \AdaBracket{Q_{\datasetPolicy}(s',a') - \tau_{\mathcal{D}} S_2(\datasetPolicy)(\cdot|s')}}. \\ 
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \frac{1}{2}\tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
    \label{eq:insample_sparsemax}
    \end{split}
\end{align}
Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \frac{\tau}{2} \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \frac{\tau}{2}$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
Under the Tsallis behavior policy assumption, the dataset support constraint $a: \datasetPolicy(a|s) > 0$ coincides with the condition $a \in K_{\mathcal{D}}(s)$.
Moreover, the assumption actually assumes a forward process of learning a set of allowable actions $K_\mathcal{D}(s)$ given $Q_{\datasetPolicy}(s,a), \forall s,a$;
 and it naturally raises the possibility of \emph{backward Tsallis learning}: given the set of allowable actions $K_\mathcal{D}(s)$, learn an optimal policy that resembles $\datasetPolicy$ in the sense of $\pi_* \preceq \datasetPolicy$, i.e. putting probability mass only on the actions in $K_\mathcal{D}$.

 The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
However, it should be noted that the learned policy no longer requires a thresholding function $\psi$ since $K_\mathcal{D}$ is given, hence no action need to be truncated.
Specifically, we propose to modify the learned policy form as 
\begin{align}
    \pi_{t+1, \datasetPolicy}(a|s) \propto \exp_2\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
    \label{eq:proposal_policy}
\end{align}
The policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.
In fact, $\pi_{t+1, \datasetPolicy}$ is a linear function in $Q_{t, \datasetPolicy}$ when $q=2$.
Another advantage is one no longer needs to compute the thresholding function $\psi$, which requires sorting in the discrete control setting; it remains unclear how to accurately estimate it for the continuous control problems.

% That is, the backward Bellman optimality equation becomes:
% \begin{align}
%     \begin{split}
%         \datasetOptimalQ(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')}}. 
%     \end{split}
% \end{align}
% The term inside the expectation is now:
% \begin{align}
%     \begin{split}
%         & \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')} \\ 
%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') + \frac{\tau}{2}(1 - \pi_{*, \datasetPolicy}(a'|s')) } \\
%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') - \frac{\tau}{2}\pi_{*, \datasetPolicy}(a'|s') } + \frac{\tau}{2} \\
%          &= \sum_{a' \in K_\mathcal{D}(s')} \frac{\AdaRectBracket{ \tau^{-1}Q_{t, \datasetPolicy}(s,a)}_+ }{\sum_{b\in K_\mathcal{D}(s')}\tau^{-1} Q_{t, \datasetPolicy}(s,b)}
%     \end{split}
% \end{align}
\textbf{Difference with In-sample Softmax. }
Our proposal is seemingly similar to in-sample softmax \cite{Xiao2023-InSampleSoftmax}.
However, the two methods are radically different: in-sample softmax requires satisfying the dataset support constraint during learning to alleviate the out-of-distribution error, and is motivated by the observation that taking softmax is easier than the hard max operator \eq{\ref{eq:hardmax_offline}}.
On the other hand, we assume the behavior policies are Tsallis policies and as such the actions not present in the dataset are those being truncated.
This viewpoint leads to the natural satisfaction of the dataset support constraint.

@@

{\color{red}(state KL divergence result here if available)}

@@

We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart for $q=2$.
In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
\begin{align}
    \begin{split}
        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_2{Q_{\theta}(s,a)} \\
        &= \datasetPolicy(a|s)  \exp_2{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_2{Q_{\theta}(s,a)}\\
        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
        \label{eq:tsallis_inac_policy}
    \end{split}
\end{align}
In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_2$ function.

@@

We now summarize the loss functions for implementing the Tsallis in-sample actor-critic algorithm:
\begin{align}
    % & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
    % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_2\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
    & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
    & \mathcal{L}_\text{critic}(\theta) = \expectation{s,a,r,s' \sim \mathcal{D}}{\AdaBracket{r + \gamma v_{\zeta}(s') - Q_{\theta}(s,a)}^2 }.
\end{align}

>>>>

A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
\begin{align}
    \begin{split}
        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
    \end{split}
\end{align}
When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
\begin{align*}
    &\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau } - \tilde{\psi}\AdaBracket{\frac{Q_k(s, \cdot)}{\tau }}}, \\
    &{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|}, \,\, \tilde{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} := \psi\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} + \frac{1}{q-1}
\end{align*}
where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.245\textwidth}
        \includegraphics[width=\textwidth]{img/q2_Gaussian_sparsemax.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.245\textwidth}
        \includegraphics[width=\textwidth]{img/q50_scaled_Gaussian_sparsemax.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.245\textwidth}
        \includegraphics[width=\textwidth]{img/q2.0_Boltzmann_sparsemax.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.245\textwidth}
        \includegraphics[width=\textwidth]{img/q50_Boltzmann_sparsemax.png}
    \end{subfigure}
    \caption{The sparsemax operator acting upon Gaussian and Boltzmann policies for $q=2$ and $q=50$ by truncating actions with values lower than $\psi$. 
    As $q\rightarrow \infty$ the policy tends toward uniform distribution.
    We assume the The dataset is generated by Tsallis policies such that the actions absent from the offline dataset are those being truncated.
    }
\end{figure}

@@

The key to both in-sample max \cite{Fujimoto2019-InSampleMax} and in-sample softmax \cite{Xiao2023-InSampleSoftmax} is the dataset support constraint.
In this paper, we propose to exploit the properties of sparsemax policies to naturally satisfy this constraint.
In a nutshell, we assume the behavior policies are Tsallis sparsemax policies, and therefore the actions not present in the dataset are with low probability and are exactly these truncated actions.
For simplicity, suppose the dataset is generated by the Tsallis sparsemax policy:
\begin{align}
    \datasetPolicy(a|s) = \AdaRectBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \psi\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}_{+},  \quad \sum_{a\in K_\mathcal{D}(s)} \pi_\mathcal{D}(a|s)= 1,
\end{align}
where $\tau_\mathcal{D}$ is a unknown coefficient and $K_\mathcal{D}(s)$ denotes the set of allowable actions.
Now let us inspect the Bellman equation for the behavior policy:
\begin{align}
    \begin{split}
    Q_{\datasetPolicy}(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a'} \datasetPolicy(a'|s') \AdaBracket{Q_{\datasetPolicy}(s',a') - \tau_{\mathcal{D}} S_2(\datasetPolicy)(\cdot|s')}}. \\ 
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a' \in K_{\mathcal{D}}(s')} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}\\
    &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \tau_{\mathcal{D}} \AdaBracket{ \sum_{a': \datasetPolicy(a'|s') > 0} \AdaBracket{\frac{Q_{\datasetPolicy}(s',a')}{\tau_{\mathcal{D}}}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy}(s', \cdot)}{\tau_{\mathcal{D}}}}^2 + 1 }}.
    \label{eq:insample_sparsemax}
    \end{split}
\end{align}
Here, $\max_{\pi \in \Delta(\mathcal{A})}\pi(a|s)Q(s,a) + \tau S_2(\pi(\cdot|s)) $ attains its maximum at $V(s) = \tau \sum_{a \in K(s)}\AdaBracket{ \frac{Q(s,a)}{\tau} }^2 - \psi\AdaBracket{\frac{Q(s, \cdot)}{\tau}}^2 + \tau$ \cite{Lee2018-TsallisRAL,Martins16-sparsemax}.
It is also possible to assume that $\datasetPolicy(a|s) = \exp_q\AdaBracket{\frac{Q_{\datasetPolicy}(s,a)}{\tau_\mathcal{D}} - \tilde{\psi}\AdaBracket{\frac{Q_{\datasetPolicy}(s, \cdot)}{\tau_\mathcal{D}}}}$ for general $q>2$, though in these cases the policy does not have a closed-form expression we have to resort to approximation.
Under the Tsallis behavior policy assumption, the dataset support constraint $a: \datasetPolicy(a|s) > 0$ coincides with the condition $a \in K_{\mathcal{D}}(s)$.
Moreover, the assumption actually assumes a forward process of learning a set of allowable actions $K_\mathcal{D}(s)$ given $Q_{\datasetPolicy}(s,a), \forall s,a$;
 and it naturally raises the possibility of \emph{backward Tsallis learning}: given the set of allowable actions $K_\mathcal{D}(s)$, learn an optimal policy that resembles $\datasetPolicy$ in the sense of $\pi_* \preceq \datasetPolicy$, i.e. putting probability mass only on the actions in $K_\mathcal{D}$.

 The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
% However, it should be noted that the learned policy no longer requires a thresholding function $\tilde{\psi}$ since the set of allowable actions $K_\mathcal{D}$ is given, hence no action need to be explicitly truncated.
% Specifically, we propose to modify the learned policy form as 
\begin{align}
    \pi_{t+1, \datasetPolicy}(a|s) = \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \tilde{\psi}{\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,\cdot)}{\tau}}}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
    \label{eq:proposal_policy}
\end{align}
When $q=2$, the policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.

% That is, the backward Bellman optimality equation becomes:
% \begin{align}
%     \begin{split}
%         \datasetOptimalQ(s,a) &= r(s,a) +  \gamma \expectation{s'\sim P(\cdot | s,a)}{ \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')}}. 
%     \end{split}
% \end{align}
% The term inside the expectation is now:
% \begin{align}
%     \begin{split}
%         & \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{\datasetOptimalQ(s',a') - \tau S_2(\pi_{*, \datasetPolicy})(\cdot|s')} \\ 
%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') + \frac{\tau}{2}(1 - \pi_{*, \datasetPolicy}(a'|s')) } \\
%          &= \sum_{a' \in K_\mathcal{D}(s')} \pi_{*, \datasetPolicy}(a'|s') \AdaBracket{ \datasetOptimalQ(s',a') - \frac{\tau}{2}\pi_{*, \datasetPolicy}(a'|s') } + \frac{\tau}{2} \\
%          &= \sum_{a' \in K_\mathcal{D}(s')} \frac{\AdaRectBracket{ \tau^{-1}Q_{t, \datasetPolicy}(s,a)}_+ }{\sum_{b\in K_\mathcal{D}(s')}\tau^{-1} Q_{t, \datasetPolicy}(s,b)}
%     \end{split}
% \end{align}

\textbf{Why use Tsallis backward learning? }
If the Tsallis dataset policy assumption holds, and Tsallis backward learning is adopted, 
then the KL divergence between the learned policy and the behavior policy can be bounded as:
\begin{align}
    \begin{split}
        &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} = \expectation{a\sim\pi_{t}(\cdot | s)}{\ln \pi_{t}(a|s) - \ln\datasetPolicy(a|s)} \\
        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
    \end{split}
\end{align}
Let us now respectively bound the three terms:
\begin{align}
    \begin{split}
        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
        % &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}  + \ln\pi_t(a|s) \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}
        & \leq (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} - \ln\pi_t (a|s)\\
        &=  \AdaBracket{ (q-1)\pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{1}{q-1},
    \end{split}
\end{align}
where we leveraged the fact that $-\frac{1}{q-1} < \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} < 0, \, \forall a\in K(s)$. If $a \notin K_{\mathcal{D}}(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
We repeat the same trick to yield an upper bound $\AdaBracket{ (q-1)\datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \frac{1}{q-1}$  for (3); and and upper bound $\frac{1}{q-1}$  for (2).

As a summary, we can state for $q>1$ that:
\begin{align}
    \begin{split}
    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  \\
    &\quad K_{\mathcal{D}}(s){\AdaRectBracket{\AdaBracket{ (q-1) \datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \AdaBracket{  (q-1) \pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{3}{q-1}}}.
\end{split}
\end{align}
This result suggests that in order to keep the learned policy close to the behavior policy, one should choose $q$ such that $(q-1)\pi^{q-1}_t(a|s)$ close to 1.

\textbf{Difference with In-sample Softmax. }
Our proposal is seemingly similar to in-sample softmax \cite{Xiao2023-InSampleSoftmax}.
However, the two methods are radically different: in-sample softmax requires satisfying the dataset support constraint during learning to alleviate the out-of-distribution error, and is motivated by the observation that taking softmax is easier than the hard max operator \eq{\ref{eq:hardmax_offline}}.
On the other hand, we assume the behavior policies are Tsallis policies and as such the actions not present in the dataset are those being truncated.
This viewpoint leads to the natural satisfaction of the dataset support constraint.

@@

@@

We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart.
In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
We found it improves the performance by simply removing $\psi$ and renormalize the policy.
In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
\begin{align}
    \begin{split}
        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
        \label{eq:tsallis_inac_policy}
    \end{split}
\end{align}
In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.

@@

We now summarize the loss functions for implementing the Tsallis in-sample actor-critic algorithm:
\begin{align}
    % & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
    % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
    & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
    & \mathcal{L}_\text{critic}(\theta) = \expectation{s,a,r,s' \sim \mathcal{D}}{\AdaBracket{r + \gamma v_{\zeta}(s') - Q_{\theta}(s,a)}^2 }.
\end{align}
